### 2012-12-11 hdf5/discs_2012/background.tex
For camera ready version

Virtual Object Layer (VOL):
VOL is a new abstraction layer internal to the HDF5 library~\cite{vol}. It is implemented just below the public API. The VOL exports an interface that allows writing plugins for HDF5, thereby enabling developers to store objects in a format different from the default HDF5 file format (like native netCDF or HDF4 format). Plugin writers provide an implementation for a set of functions that access data on disk. These include functions for file management, dataset creation and access, group creation, to name a few.

@@

>>>>

Virtual Object Layer (VOL):
VOL is a new abstraction layer internal to the HDF5 library~\cite{vol}. As shown in figure~\ref{hdf5_vol} it is implemented just below the public API. The VOL exports an interface that allows writing plugins for HDF5, thereby enabling developers to store objects in a format different from the default HDF5 file format (like native netCDF or HDF4 format). Plugin writers provide an implementation for a set of functions that access data on disk. These include functions for file management, dataset creation and access, group creation, to name a few.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in,height=2.8in]{hdf5_vol}
\caption{Virtual Object Layer (VOL)}
\label{hdf5_vol}
\end{figure}

@@

As stated before, when multiple processes access a single, shared file, file systems like Lustre are known to exhibit sub-standard I/O performance. PLFS has demonstrated benefits of converting this N-1 access pattern into one where every process writes data to its own physical file and sufficient metadata is maintained to recreate the original file in the correct order of writes issued. 
Secondly, the data model of HDF5 stores multiple datasets in a single file on disk. As a result, one cannot perform any useful post-processing on individual HDF5 objects. As we shall see, storing objects in a different file format allows us to perform optimizations on individual HDF5 objects without affecting the remaining data in the file.

### 2012-12-11 hdf5/discs_2012/evaluation.tex
For camera ready version

For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool~\cite{h5perf}. It allows configuring various parameters, such as the number of processes, number of datasets, amount of data read/written by a process in a single I/O call (transfer size) etc. For our measurements, we created 10 datasets, and the total file size was 64 GB or more.
In every run, every process contributed equal amount of data per dataset using the default individual (non-collective) I/O mode of h5perf.

>>>>

In this section, We present the the I/O performance of our plugin. These experiments are not targeted towards performing a semantic analysis of data, such as active analysis or semantic restructuring discussed in the previous section.
For evaluation purposes, we have used HDF5's \textit{h5perf} performance tool~\cite{h5perf}. It allows configuring various parameters, such as the number of processes, number of datasets, amount of data read/written by a process in a single I/O call (transfer size) etc. For our measurements, we created 10 1-dimensional datasets, and the total file size was 64 GB or more.
In every run, every process contributed equal amount of data per dataset using the default individual (non-collective) I/O mode of h5perf.

### 2012-12-11 hdf5/discs_2012/introduction.tex
For camera ready version

An HDF5 file is a self-describing format which combines data and metadata. It is a container in which users typically store multiple HDF5 objects alongside their metadata. 
However, this native single-file format has its disadvantages. A PHDF5 application has multiple processes accessing a single file (i.e. N-1 access pattern). Unfortunately, many popular parallel file systems are known to behave poorly under these circumstances~\cite{lustre_perf,nersc_io}. Secondly, since many HDF5 objects are stored in a single file, this complicates performing useful semantic operations on individual objects.

>>>>

An HDF5 file is a self-describing format which combines data and metadata. It is a container in which users typically store multiple HDF5 objects alongside their metadata. 
A PHDF5 application has multiple processes accessing a single file (i.e. N-1 access pattern). Unfortunately, many popular parallel file systems are known to behave poorly under these circumstances~\cite{lustre_perf,nersc_io}. Secondly, HDF5 stores multiple data objects in a single file. This complicates performing useful semantic operations on individual objects.

### 2012-12-11 hdf5/discs_2012/plugin.tex
For camera ready version

We have developed a plugin which stores data in a format which enables performing semantic analysis on the data and uses PLFS for high I/O throughput. 
Instead of storing all objects in a single file, it stores every HDF5 object in a separate location, so that data from different objects is not contained in the same file. In short, we provide a raw mapping of HDF5 objects to the file system. 
HDF5 files and groups are stored as directories, whereas datasets and attributes, which contain raw data, are stored in files created using PLFS. 
Consider the example shown in figure~\ref{hdf5_example}. Using the plugin, file Sample.h5 is stored as a directory. Group G1 is stored as a directory under it, and datasets D1 and D2 are files. Attribute A1 is a file created at the same location as dataset D1 to which it is attached. Additionally, the name of the attribute is stored as \textit{dataset name}.\textit{attribute name} to denote the dataset to which the attribute is attached. Thus these objects are stored at the following paths:\\
/Sample.h5/\\
/Sample.h5/G1\\
/Sample.h5/G1/D3\\
/Sample.h5/D1\\
/Sample.h5/D1.A1\\
/Sample.h5/D2\\

@@

%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{hdf5_example_plugin}
%\caption{HDF5 file created using our PLFS plugin}
%\label{hdf5_example_plugin}
%\end{figure}

@@

Active analysis borrows the transducers idea from the Semantic File System~\cite{semantic_fs} which has since been productized in Google's BigTable and Apache Hbase technologies~\cite{google_coprocessors,GFS,apache_hbase}.  With active analysis, the application can ship a data parser function when it creates the PLFS file.  As the data is written into PLFS, PLFS can apply the data function on the streaming data.  The function will output key-value pairs which PLFS can embed in its extensible metadata.  In this example, one simple function might record the height of the largest wave.  Due to PLFS's model of storing a logical file across multiple physical files, the PLFS extensible metadata can record the height of the largest wave within each physical file.  However, given that PLFS now understands the structure of the logical file, these multiple physical files, within the PLFS container, are more accurately thought of as shards.  In a future burst buffer architecture~\cite{burst_buffers}, these semantic shards will be spread across multiple burst buffer nodes.  Therefore, subsequent analysis of the ocean model can quickly find the burst buffer containing the shard with the largest wave by searching a small amount of extensible metadata instead of scanning the entire ocean model.

Semantic restructuring is the idea of reorganizing the data into a new set of semantic shards.  This would be done to speed future analysis routines.  For example, assume that the ocean model was originally sharded using a row-order organization (i.e. across latitude instead of longitude).  An analysis routine which will explore the model along a column-ordering will suffer poor performance with the row-order organization as its access pattern will result in a large number of small reads from a large set of semantic shards.  However, by knowing the semantic structure, the analysis routine can request a semantic restructuring which will be a compact, intuitively described request such as "restructure into row-ordering."  Without structural knowledge, a semantic restructuring would be significantly more complicated: the analysis routine would have to send a large list of logical offsets to PLFS to inform it of expected read patterns.  In an exascale system, the list of logical offsets will be in the order of one billion.  Semantic restructuring shrinks the size of the request to a small constant value.

>>>>

We have developed a plugin which stores data in a format which enables performing semantic analysis on the data and uses PLFS for high I/O throughput. 
Instead of storing all objects in a single file, it stores every HDF5 object in a separate location, so that data from different objects is not contained in the same file. In short, we provide a raw mapping of HDF5 objects to the file system. 
HDF5 files and groups are stored as directories, whereas datasets and attributes, which contain raw data, are stored in files created using PLFS. 
Consider the example shown in figure~\ref{hdf5_example}. Using the plugin, file Sample.h5 is stored as a directory. Group G1 is stored as a directory under it, and datasets D1 and D2 are files. Attribute A1 is a file created at the same location as dataset D1 to which it is attached. Additionally, the name of the attribute is stored as \textit{dataset name}.\textit{attribute name} to denote the dataset to which the attribute is attached. Thus these objects are stored at the following paths as shown in figure ~\ref{hdf5_example_plugin}:\\
/Sample.h5/\\
/Sample.h5/G1\\
/Sample.h5/G1/D3\\
/Sample.h5/D1\\
/Sample.h5/D1.A1\\
/Sample.h5/D2\\

@@

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{hdf5_example_plugin}
\caption{HDF5 data stored using the plugin}
\label{hdf5_example_plugin}
\end{figure}

@@

Active analysis borrows the transducers idea from the Semantic File System~\cite{semantic_fs} which has since been productized in Google's BigTable and Apache Hbase technologies~\cite{google_coprocessors,GFS,apache_hbase}.  With active analysis, the application can ship a data parser function when it creates the PLFS file.  As the data is written into PLFS, PLFS can apply the data function on the streaming data.  The function will output key-value pairs which PLFS can embed in its extensible metadata (figure~\ref{active_analysis}).  In this example, one simple function might record the height of the largest wave.  Due to PLFS's model of storing a logical file across multiple physical files, the PLFS extensible metadata can record the height of the largest wave within each physical file.  However, given that PLFS now understands the structure of the logical file, these multiple physical files, within the PLFS container, are more accurately thought of as shards.  In a future burst buffer architecture~\cite{burst_buffers}, these semantic shards will be spread across multiple burst buffer nodes.  Therefore, subsequent analysis of the ocean model can quickly find the burst buffer containing the shard with the largest wave by searching a small amount of extensible metadata instead of scanning the entire ocean model.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in,height=3.0in]{active_analysis}
\caption{Active Analysis of Data}
\label{active_analysis}
\end{figure}

Semantic restructuring is the idea of reorganizing the data into a new set of semantic shards.  This would be done to speed future analysis routines.  For example, assume that the ocean model was originally sharded using a row-order organization (i.e. across latitude instead of longitude).  An analysis routine which will explore the model along a column-ordering will suffer poor performance with the row-order organization as its access pattern will result in a large number of small reads from a large set of semantic shards.  However, by knowing the semantic structure, the analysis routine can request a semantic restructuring (figure~\ref{semantic_restructuring}) which will be a compact, intuitively described request such as "restructure into column-ordering."  Without structural knowledge, a semantic restructuring would be significantly more complicated: the analysis routine would have to send a large list of logical offsets to PLFS to inform it of expected read patterns.  In an exascale system, the list of logical offsets will be in the order of one billion.  Semantic restructuring shrinks the size of the request to a small constant value. A storage format where datasets are stored in separate files allows us to make modifications to the storage layout of one dataset without affecting other datasets and data.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in,height=3.0in]{semantic_restructuring}
\caption{Semantic Restructuring of Data}
\label{semantic_restructuring}
\end{figure}
